---
title: "Assignment03"
author: "Shan"
date: "10/13/2020"
output: github_document
---

```{r,setup}
knitr::opts_chunk$set(echo = TRUE)
library(xml2)
library(httr)
library(stringr)
library(readr)
library(tidytext)
library(dplyr)
library(tidyverse)
library(ggplot2)
```
#Part1 APIs
Using the NCBI API, look for papers that show up under the term "sars-cov-2 trial vaccine." Look for the data in the pubmed database, and then retrieve the details of the paper as shown in lab 7. How many papers were you able to find?

Using the list of pubmed ids you retrieved, download each papers’ details using the query parameter rettype = abstract. If you get more than 250 ids, just keep the first 250.

As we did in lab 7. Create a dataset containing the following:
Pubmed ID number,
Title of the paper,
Name of the journal where it was published,
Publication date, and
Abstract of the paper (if any).

#count the number of paper 
```{r}
# Downloading the website
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2+trial+vaccine")
# Finding the counts
counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/div[2]/div[1]/span")
# Turning it into text
counts <- as.character(counts)
# Extracting the data using regex
stringr::str_extract(counts, "[0-9,]+")
```
-There are 560 papers about "sars-cov-2 trial vaccine" in the pubmed.
#covid-trial-vaccine
```{r}
library(httr)
query_ids <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(
               db = "pubmed",
               term = "sars-cov-2 trial vaccine",
               retmax =  1000
            )
)
```

#get ids
```{r}
# Extracting the content of the response of GET
ids <- httr::content(query_ids)
# Turn the result into a character vector
ids <- as.character(ids)
# Find all the ids 
ids <- stringr::str_extract_all(ids, "<Id>[0-9]+</Id>")[[1]]
# Remove all the leading and trailing <Id> </Id>. Make use of "|"
ids <- stringr::str_remove_all(ids, "</?Id>")
```
#get abstracts
```{r}
publications <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi",
  query = list(
    db = "pubmed",
    id = paste(ids,collapse = ","),
    retmex = 1000,
    rettype = "abstract"
    )
)

# Turning the output into character vector
publications <- httr::content(publications)
publications_txt <- as.character(publications)
```
#form a dataset
```{r}
pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)
```
Now, extract the abstract and article title for each one of the elements of
`pub_char_list`. You can either use `sapply()` as we just did, or simply
take advantage of vectorization of `stringr::str_extract`
#titles journals,etc
```{r}
# titles
titles <- str_extract(pub_char_list, "<ArticleTitle>(\\n|.)+</ArticleTitle>")
titles <- str_remove_all(titles, "</?[[:alnum:]]+>")
titles <- str_replace_all(titles, "\\s+", " ")
#  journal names
journals <- str_extract(pub_char_list, "<Title>(\\n|.)+</Title>")
journals <- str_remove_all(journals, "</?[[:alnum:]]+>")
journals <- str_replace_all(journals, "\\s+", " ")
#  publication date
dates <- str_extract_all(pub_char_list, "<PubDate>(\\n|.)+</PubDate>")
dates <- str_remove_all(dates, "</?[[:alnum:]]+>")
dates <- str_replace_all(dates, "\\s+", " ")
#  abstracts 
abstracts <- str_extract(pub_char_list, "<Abstract>(\\n|.)+</Abstract>")
abstracts <- str_remove_all(abstracts, "</?[[:alnum:]]+>")
abstracts <- str_replace_all(abstracts, "\\s+", " ")
table(is.na(abstracts))
```
#build the database
```{r}
database <- data.frame(
  PubMedID = ids,
  Title = titles,
  Journal = journals,
  Date = dates,
  Abstract = abstracts
)
knitr::kable(database)
```

#Part 2 Text Mining 
A new dataset has been added to the data science data repository "https://github.com/USCbiostats/data-science-data/tree/master/03_pubmed."The dataset contains 3241 abstracts from articles across 5 search terms. Your job is to analyse these abstracts to find interesting insights.
# read in the file
```{r}
mydata <- read_csv("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/03_pubmed/pubmed.csv")
head(mydata)
```
#Tokenize the abstracts&count the number
```{r}
mydata%>%
  unnest_tokens(output = token,input = abstract)%>%
  count(token, sort = TRUE)
```
#remove the stop words
```{r}
mydata %>%
  unnest_tokens(token,abstract)%>%
  anti_join(stop_words, by = c("token" = "word"))%>%
  count(token,sort = TRUE)
```
#get the top 5 by terms
```{r}
mydata %>%
  unnest_tokens(token,abstract)%>%
  anti_join(stop_words, by = c("token" = "word"))%>%
  group_by(term)%>%
  count(token)%>%
  top_n(5,n)
```

-8 out of 10 top showed most word in the article are stop words.Removing the stop words does change the frequency of tokens.The top 5 common tokens for each search terms are shown above. 

#tokenize the abstracts into bigrams
```{r}
mydata %>%
  unnest_ngrams(token, abstract, n=2) %>%
  count(token, sort = TRUE) %>%
  top_n(10, n)%>%
  ggplot(aes(x = n, y = fct_reorder(token, n)))+
  geom_col()
```
#calculate the TF-IDF value
```{r}
mydata %>%
  unnest_tokens(token, abstract) %>%
  group_by(term)%>%
  anti_join(stop_words, by = c("token" = "word"))%>%
  count(token, sort = TRUE)%>%
  bind_tf_idf(token, term, n) %>%
  top_n(5,n)%>%
  arrange(desc(tf_idf))
```
  
What are the 5 tokens from each search term with the highest TF-IDF value? How are the results different from the answers you got in question 1?
covid:"pandemic" has the highest tf-idf value;
prostate cancer: "postate";
preeclampsia:"eclampsia";
meningitis:"meningitis";
cystic fibrosis:"cf"

